---
title: "Exercises I: Statistical Modeling in R"
author: "Henrik Singmann"
date: "25 July 2018"
output: pdf_document
---


```{r setup, include=FALSE}
require(psych)
data(sat.act)
sat.act$gender <- factor(sat.act$gender, 1:2, labels = c("male", "female"))
sat.act$education <- factor(sat.act$education)
sat.act <- na.omit(sat.act)
```

# Formula Interface for Statistical Models: `~`

- Allows symbolic specification of statistical model, e.g. linear models: `lm(ACT ~ SATQ, sat.act)`
- Everything to the left of `~` is the dependent variable.
- Independent variables are to the right of the `~`:

| Formula | &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; | Interpretation  |
| ------------------------|---|----------------------------------|
| `~ x` or `~1+x`         || Intercept and main effect of `x` | 
| ` ~ x-1` or `~0 + x`    || Only main effect of `x` and no intercept (questionable) |
| `~ x+y`                 || Main effects of `x` and `y`|
| `~ x:y`                 || Interaction between `x` and `y` (and no main effect) |
| `~ x*y` or `~ x+y+x:y`  || Main effects and interaction between `x` and `y` |


# Continuous Variables: How many Parameters in each Model?

```{r, eval=FALSE}
lm(ACT ~ SATQ_c + SATV_c, sat.act)   # a
lm(ACT ~ SATQ_c : SATV_c, sat.act)   # b
lm(ACT ~ 0 + SATQ_c:SATV_c, sat.act) # c
lm(ACT ~ SATQ_c*SATV_c, sat.act)     # d
lm(ACT ~ 0+SATQ_c*SATV_c, sat.act)   # e
```

# Categorical Variables: How many Parameters in each Model?

```{r, eval=FALSE}
lm(ACT ~ gender, sat.act)                  # a
lm(ACT ~ 0+gender, sat.act)                # b
lm(ACT ~ gender+education, sat.act)        # c
lm(ACT ~ 0+gender+education, sat.act)      # d
lm(ACT ~ gender:education, sat.act)        # e
lm(ACT ~ 0+gender:education, sat.act)      # f
lm(ACT ~ gender*education, sat.act)        # g
lm(ACT ~ 0+gender*education, sat.act)      # h
lm(ACT ~ gender+gender:education, sat.act) # i
```

```{r}
levels(sat.act$gender) ## 2
levels(sat.act$education) ## 6
```
