---
title: "Exercise 2: Identifying Random Effects-Structure"
author: "Henrik Singmann"
date: "25 July 2018"
output: pdf_document
---

# Exercise 2: Identifying Random Effects-Structure

Your task is to identify the *maximal random-effects structure justified by the design* (Barr, Levy, Scheepers, & Tily, 2013) for one data set and implement this structure in `lme4::lmer` syntax.


# Freeman, Heathcote, Chalmers, and Hockley (2010)

Lexical decision and word naming latencies for 300 words and 300 nonwords presented in Freeman, Heathcote, Chalmers, and Hockley (2010). The study had one between-subjects factors, `task` with two levels (`"naming"` or `"lexdec"`), and four within-subjects factors: `stimulus` type with two levels (`"word"` or `"nonword"`), word `density` and word `frequency` each with two levels (`"low"` and `"high"`) and stimulus `length` with three levels (`4`, `5`, and `6`).

The data comes with `afex` as `fhch2010`:
```{r}
data("fhch2010", package = "afex")
str(fhch2010)
```

What is the maximal random-effects structure justified by the design for this data set for the dependent variable `log_rt`:

```{r, eval=FALSE}
mixed(log_rt ~ ...)

```


## References 
- Barr, D. J., Levy, R., Scheepers, C., & Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: Keep it maximal. *Journal of Memory and Language*, 68(3), 255-278. https://doi.org/10.1016/j.jml.2012.11.001
-Freeman, E., Heathcote, A., Chalmers, K., & Hockley, W. (2010). Item effects in recognition memory for words. *Journal of Memory and Language*, 62(1), 1-18. http://doi.org/10.1016/j.jml.2009.09.004
