---
title: "Exercise 2"
output: html_notebook
---

# Exercise 2: Identifying Random Effects-Structure

Your task is to identify the *maximal random-effects structure justified by the design* (Barr, Levy, Scheepers, & Tily, 2013) for two data sets and implement this structure in `lme4::lmer` syntax.

## Data-Set 1: Skovgaard-Olsen et al. (2016): The Relevance Effect and Conditionals

- Conditional = *if-then* statement; e.g., If global warning continues, London will be flooded.
- Bayesian reasoning often assumes 'the Equation': *P*(if *A* then *C*) = *P*(*C*|*A*)
- Our question: Does the Equation hold even if no apparent relationship between *A* and *C*? 
  - positive relevance (PO): *A* is a reason for *C* 
  - negative relevance (NE): *A* is a reason against *C* 
  - irrelevance (IR): *A* and *C* have no apparent relationship 

- 183 (of 348) participants recruited via `crowdflower.com` worked on 12 items.

Full design for participants: 4 x 3 x 2. First two factors repeated-measures, last factor between-subjects.

Each participant worked on 12 items, 4 estimates of *P*(*C*|*A*) (`c_given_a`) for each of 3 relevance conditions (`rel_cond`):

- positive relavnce (PO): *A* is a reason for *C* (IF Sophia buys an orc costume for Tim, THEN Tim will be excited about his present.)
- negative relevance (NE): *A* is a reason against *C* (IF Sophia buys a Barbie doll for Tim, THEN Tim will be excited about his present.) 
- irrelevance (IR): *A* and *C* have no apparent relationship (IF Sophia regularly wears shoes, THEN Tim will be excited about his present.)

Type of DV (`dv_question`) is between-subjects condition: 

- Probability of 'if *A* then *C*': How probable is 'if *A* then *C*'?
- Acceptability of 'if *A* then *C*': How acceptable is it to say 'if *A* then *C*'?).

In total we created 12 vignettes (i.e., items). Each participant provided one responses for each vignette with the condition of the vignette randomly chosen among the 4 x 3 within-subjects design.

The data is given in:

```{r}
load("ssk16_dat_preapred.rda") # data preapred in 'prepare_data.R'
str(dat)
```


What is the `lme4::lmer` or `afex::mixed` formula for the maximal random-effects structure for this design?

```{r, eval=FALSE}
mixed(dv ~ ...)
```

## Data-Set 2: Freeman, Heathcote, Chalmers, and Hockley (2010)

Lexical decision and word naming latencies for 300 words and 300 nonwords presented in Freeman, Heathcote, Chalmers, and Hockley (2010). The study had one between-subjects factors, `task` with two levels (`"naming"` or `"lexdec"`), and four within-subjects factors: `stimulus` type with two levels (`"word"` or `"nonword"`), word `density` and word `frequency` each with two levels (`"low"` and `"high"`) and stimulus `length` with three levels (`4`, `5`, and `6`).

The data comes with `afex` as `fhch2010`:
```{r}
data("fhch2010", package = "afex")
str(fhch2010)
```

What is the maximal random-effects structure justified by the design for this data set for the dependent variable `log_rt`:

```{r, eval=FALSE}
mixed(log_rt ~ ...)

```


## References 
- Barr, D. J., Levy, R., Scheepers, C., & Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: Keep it maximal. *Journal of Memory and Language*, 68(3), 255-278. https://doi.org/10.1016/j.jml.2012.11.001
-Freeman, E., Heathcote, A., Chalmers, K., & Hockley, W. (2010). Item effects in recognition memory for words. *Journal of Memory and Language*, 62(1), 1-18. http://doi.org/10.1016/j.jml.2009.09.004
- Skovgaard-Olsen, N., Singmann, H., & Klauer, K. C. (2016). The relevance effect and conditionals. *Cognition*, 150, 26-36. https://doi.org/10.1016/j.cognition.2015.12.017

